grad_weight

### 2022-04-12 inst/grad_weight.Rmd
bug fixed: differentiate WRT w, not D

---
title: "grad_weight"
author: "Robin KS Hankin"
date: "10/04/2022"
output: html_document
---

@@

```{r,label=definegws}
grad_weight_single <- function(H, probs, wpv, wanted){
    jj <- H[wpv]
    alpha <- drop(powers(jj))
    powers(jj) <- 1
print(wanted)
    alpha*sum(probs[names(probs) %in% wanted])/loglik(probs,jj,log=FALSE)
}
```

@@

```{r}
grad_weight_single

@@

brackfunc <- function(w,D){
  list(
  b1 = c(p1=w),
  b2 = c(p1=D, p2=D),	
  b3 = c(p2=1),
  b4 = c(p1=w+D,p2=1+D),
  b5 = c(p2=w),
  b6 = c(p1=1),
  b7 = c(p1=1+D,p2=w+D),
  b8 = c(p3=w),
  b9 = c(p3=w,p1=1)
)
}
brackfunc(w,D)

wanted <- list(NULL,c("p1","p2"),NULL,c("p1","p2"), NULL,NULL,c("p1","p2"),NULL,NULL)
wanted

@@

Differentiate with respect to `D`:

```{r}
diff <- 0
dw <- 1e-4
probs <- c(p1=0.15 , p2=0.45, p3=0.4)
B <- brackfunc(w,D)
for(i in seq_along(wanted)){
                                   diff <- diff +  grad_weight_single(H, probs, wpv=B[[i]],wanted=wanted[[i]])
                         print(grad_weight_single(H, probs, wpv=B[[i]],wanted=wanted[[i]]))

}

LHS <- diff
RHS <-  (loglik(probs,f2(w=w+dw,D=D)) - loglik(probs,f2(w=w,D=D)))/dw  # differential via numerics
c(LHS,RHS,LHS-RHS)
```

>>>>

<---
title: "grad_weight"
author: "Robin KS Hankin"
date: "10/04/2022"
output: html_document
---

@@

```{r,label=definegws}
grad_weight_single <- function(H, probs, wpv, wanted){
    jj <- H[wpv]
    alpha <- drop(powers(jj))
    powers(jj) <- 1
    alpha*sum(probs[names(probs) %in% wanted])/loglik(probs,jj,log=FALSE)
}
```

@@

```{r}

@@

brackfunc <- function(w,D){
  list(
  b1 = c(p1=w),
  b2 = c(p1=D, p2=D),	
  b3 = c(p2=1),
  b4 = c(p1=w+D,p2=1+D),
  b5 = c(p2=w),
  b6 = c(p1=1),
  b7 = c(p1=1+D,p2=w+D),
  b8 = c(p3=w),
  b9 = c(p3=w,p1=1)
)
}

wanted <- list("p1",NULL,NULL,"p1","p2",NULL,"p2","p3","p3")
wanted

@@

Differentiate with respect to `w`:

```{r}
diff <- 0
dw <- 1e-4
probs <- c(p1=0.15 , p2=0.45, p3=0.4)
B <- brackfunc(w,D)
for(i in seq_along(wanted)){
   diff <- diff +  grad_weight_single(H, probs, wpv=B[[i]],wanted=wanted[[i]])
}

LHS <- diff  # using grad_weight_single()
RHS <- (loglik(probs,f2(w=w+dw,D=D)) - loglik(probs,f2(w=w,D=D)))/dw  # differential via numerics
c(LHS,RHS,LHS-RHS)
```

We may even be a little clever  and use a central difference:

```{r}
RHS2 <- (loglik(probs,f2(w=w+dw/2,D=D)) - loglik(probs,f2(w=w-dw/2,D=D)))/dw  # differential via numerics
c(LHS,RHS2,LHS-RHS2)
```
